---
title: "chapter3.rmd"
author: "Ilse Ekman"
date: "18 11 2019"
output: html_document
---

### Exercise 3


## Analysis of the third exercise

```{r}
alc <-read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep = ",", header = TRUE)
```


**Structure of the data:**

```{r}
str(alc)
```



This data is a combined data from two datasets regarding the performance in mathematics and Portuguese language.

The data has 382 observations of 35 variables.

**The names of the variables:**

```{r}
colnames(alc)
```

**The purpose of my analysis is to study the relationship between alcohol consumption and following four variables:**

**1.sex: I hypothesize that male students have higher consumption of alcohol.**

**2.age: I hypothesize that the older the student is the more they consume alcohol.**

**3.goout: I hypothesize that the more the student goes out with friends the bigger their alcohol consumption is.**

**4.health: I hypothesize that the current health status is worse in the students with higher alcohol consumption.**

**Exploring the first hypothesis:**

```{r}
library(dplyr); library(ggplot2)

alc %>% group_by(sex, high_use) %>% summarise(count = n())

g1 <- ggplot(alc, aes(x = sex, y = alc_use))
g1 + geom_boxplot() + ylab("Alcohol usage")
```

**1: It seems male students have higher consumption of alcohol.**

**Exploring the second hypothesis:**

```{r}
alc %>% group_by(age, high_use) %>% summarise(count = n())

g2 <- ggplot(alc, aes(x = high_use, y = age))
g2 + geom_boxplot() + ylab("Age")
```
**2: It seems older students have higher consumption of alcohol.**

**Exploring the third hypothesis:**

```{r}
alc %>% group_by(goout, high_use) %>% summarise(count = n())

g3 <- ggplot(alc, aes(x = high_use, y = goout))
g3 + geom_boxplot() + ylab("Going out")
```
**3: Going out more frequently seems to correlate to higher alcohol consumption.**

**Exploring the fourth hypothesis:**

```{r}
alc %>% group_by(health, high_use) %>% summarise(count = n())

g4 <- ggplot(alc, aes(x = high_use, y = health))
g4 + geom_boxplot() + ylab("Health")
```
**4: Current health status of a student does not seem to reflect the alcohol consumption.**

**Exploring the variables with logistic regression:**

```{r}
m <- glm(high_use ~ sex + age + goout + health, data = alc, family = "binomial")

summary(m)
coef(m)
```

**To compute out odds ratios:**

```{r}
OR <- coef(m) %>% exp
```

**To compute out confidence intervals:**

```{r}
CI <- confint(m) %>% exp

cbind(OR, CI)
```

**To predict the probability of high alcohol usage:**

probabilities <- predict(m, type = "response")

alc <- mutate(alc, probability = probabilities)

alc <- mutate(alc, prediction = probability > 0.5)

select(alc, failures, absences, sex, high_use, probability, prediction) %>% tail(10)

table(high_use = alc$high_use, prediction = alc$prediction)