---
title: "chapter6.Rmd"
author: "Ilse Ekman"
date: "8 12 2019"
output: html_document
---

# Chapter 6

## RATS dataset analysis

read.csv("RATSL.csv")
library(tidyr)
library(dplyr)
library(ggplot2)

### Figure 1

```{r}
p1 <- ggplot(RATSL, aes(x = Time, y = Weight, group = ID))
p2 <- p1 + geom_text(aes(label = Group))
p3 <- p2 + scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10))
p4 <- p3 + scale_y_continuous(name = "Weight (grams)")
p5 <- p4 + theme_bw()
p6 <- p5 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p6
```

### Standardization and Figure 2

RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate( stdweight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()
  
```{r}
glimpse(RATSL)
```

```{r}
p1 <- ggplot(RATSL, aes(x = Time, y = stdweight, linetype = ID))
p2 <- p1 + geom_line() + scale_linetype_manual(values = rep(1:10, times=4))
p3 <- p2 + facet_grid(. ~ Group, labeller = label_both)
p4 <- p3 + theme_bw() + theme(legend.position = "none")
p5 <- p4 + theme(panel.grid.minor.y = element_blank())
p6 <- p5 + scale_y_continuous(name = "standardized weight")
p6
```

### Summary of data and Figure 3

n <- RATSL$Time %>% unique() %>% length()

RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise( mean=mean(Weight), se=sd(Weight)/sqrt(n) ) %>%
  ungroup()

```{r}
glimpse(RATSS)
```

```{r}
p1 <- ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group))
p2 <- p1 + geom_line() + scale_linetype_manual(values = c(1,2,3))
p3 <- p2 + geom_point(size=3) + scale_shape_manual(values = c(1,2,3))
p4 <- p3 + geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.5)
p5 <- p4 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p6 <- p5 + theme(legend.position = c(0.8,0.8))
p7 <- p6 + scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
p7
```

### Figure 4

```{r}
p1 <- ggplot(RATSL, aes(x = factor(Time), y = Weight, fill = Group))
p2 <- p1 + geom_boxplot(position = position_dodge(width = 0.9))
p3 <- p2 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p4 <- p3 + theme(legend.position = c(0.8,0.8))
p5 <- p4 + scale_x_discrete(name = "Time")
p5
```

### Figure 5

RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
  
```{r}
glimpse(RATSL8S)
```

```{r}
p1 <- ggplot(RATSL8S, aes(x = Group, y = mean))
p2 <- p1 + geom_boxplot()
p3 <- p2 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p4 <- p3 + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")
p5 <- p4 + scale_y_continuous(name = "mean(Weight), Time 1-64")
p5
```

### Taking out the outlier from Group 2 and Figure 6

RATSL8S1 <- RATSL8S %>%
  filter(mean < 550)
  
glimpse(RATSL8S1)
  
```{r}
p1 <- ggplot(RATSL8S1, aes(x = Group, y = mean))
p2 <- p1 + geom_boxplot()
p3 <- p2 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p4 <- p3 + stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white")
p5 <- p4 + scale_y_continuous(name = "mean(Weight), Time 1-64")
p5
```

### Anova model

```{r}
fit <- lm(mean ~ Group, data = RATSL8S1)
summary(fit)
anova(fit)
```

## BPRS dataset analysis

read.csv("BPRSL.csv")

### Figure 1

```{r}
p1 <- ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject))
p2 <- p1 + geom_line() + scale_linetype_manual(values = rep(1:10, times=4))
p3 <- p2 + facet_grid(. ~ treatment, labeller = label_both)
p4 <- p3 + theme_bw() + theme(legend.position = "none")
p5 <- p4 + theme(panel.grid.minor.y = element_blank())
p6 <- p5 + scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
p6
```

### Regression model and Figure 2

BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRS_reg)

```{r}
p1 <- ggplot(BPRSL, aes(x = week, y = bprs, group = treatment))
p2 <- p1 + geom_line(aes(linetype = treatment))
p3 <- p2 + scale_x_continuous(name = "weeks")
p4 <- p3 + scale_y_continuous(name = "bprs")
p5 <- p4 + theme_bw() + theme(legend.position = "top")
p6 <- p5 + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p6
```


### The Random Intercept Model

library(lme4)
```{r}
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref)
```

### Summary and Student t-test

BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()

```{r}
t.test(mean ~ treatment, data = BPRSL8S, var.equal = TRUE)
```

### Anova

```{r}
baseline <- BPRS$week0
BPRSL8S1 <- BPRSL8S %>%
  mutate(baseline)
  fit <- lm(mean ~ baseline + treatment, data = BPRSL8S1)
summary(fit)
anova(fit)
```

