---
title: "chapter3.rmd"
author: "Ilse Ekman"
date: "18 11 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


# Analysis


read.csv("alc.csv")

# Structure of the data:
str(alc)


# This data is a combined data from two datasets regarding the performance in mathematics and Portuguese language.
# The data has 382 observations of 35 variables.

# The names of the variables:
colnames(alc)

# The purpose of my analysis is to study the relationship between alcohol consumption and following four variables:
# 1.sex: I hypothese that male students have higher consumption of alcohol.
# 2.age: I hypothese that the older the student is the more they consume alcohol.
# 3.goout: I hypothese that the more the student goes out with friends the bigger their alcohol consumption is.
# 4.health: I hypothese that the current health status is worse in the students with higher alcohol consumption.

# Exploring the first hypothesis:
library(dplyr); library(ggplot2)

alc %>% group_by(sex, high_use) %>% summarise(count = n())

g1 <- ggplot(alc, aes(x = sex, y = alc_use))
g1 + geom_boxplot() + ylab("Alcohol usage")

# Exploring the second hypothesis:

alc %>% group_by(age, high_use) %>% summarise(count = n())

g2 <- ggplot(alc, aes(x = high_use, y = age))
g2 + geom_boxplot() + ylab("Age")

# Exploring the third hypothesis:

alc %>% group_by(goout, high_use) %>% summarise(count = n())

g3 <- ggplot(alc, aes(x = high_use, y = goout))
g3 + geom_boxplot() + ylab("Going out")

# Exploring the fourth hypothesis:

alc %>% group_by(health, high_use) %>% summarise(count = n())

g4 <- ggplot(alc, aes(x = high_use, y = health))
g4 + geom_boxplot() + ylab("Health")


# Exploring the variables with logistic regression:
m <- glm(high_use ~ sex + age + goout + health, data = alc, family = "binomial")
summary(m)
coef(m)

**To compute out odds ratios:**
OR <- coef(m) %>% exp
**To compute out confidence intervals:**
CI <- confint(m) %>% exp
cbind(OR, CI)

**To predict the probability of high alcohol usage:**
probabilities <- predict(m, type = "response")
alc <- mutate(alc, probability = probabilities)
alc <- mutate(alc, prediction = probability > 0.5)
select(alc, failures, absences, sex, high_use, probability, prediction) %>% tail(10)
table(high_use = alc$high_use, prediction = alc$prediction)